### 2021-10-28 manuscript/hiv-gwas.tex
Update on Overleaf.

\begin{figure}[H]
	\begin{center}
		\includegraphics[width=0.7\linewidth]{figures/model_figure}
		\caption{A statistical model of the spVL trait. The virus-determined part of spVL, $g_v$, evolves according to an OU process along the viral phylogeny. The dashed line shows the evolutionarily optimal trait value $\theta$. The environmentally-determined part of spVL, $e$, including direct host genetic effects and other environmental effects, is modeled as a Gaussian random variable whose value is added at the tips of the phylogeny.}
		\label{fig:spVLModel}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
		\includegraphics[width=0.75\linewidth]{figures/simulation_results}
		\caption{(A) Root mean square error (RMSE) with respect to true (simulated) values for the host part of spVL ($\bm{{g_h}}$). The left plot is for the POUMM-based estimate of the non-viral part of spVL ($\bm{\hat{e}}$) and the right plot is for spVL scaled such that it is centered at zero ($\bm{z} - \bm{\bar{z}}$). Each plot tile represents the average RMSE over 20 simulated datasets of 500 samples. (B) True positive rate (TPR) in GWAS performed using true (simulated) non-viral spVL values ($\bm{e}$), the POUMM-based estimate for the non-viral part of spVL ($\bm{\hat{e}}$), and total spVL scaled such that it is centered at zero ($\bm{z} - \bm{\bar{z}}$). Each plot tile represents the average TPR over 20 simulated datasets of 500 samples.}
		\label{fig:simulationResults}
	\end{center}
\end{figure}

>>>>

\begin{figure}[H]
	\begin{center}
		\includegraphics[width=0.75\linewidth]{figures/model_figure}
		\caption{The POUMM statistical model for the spVL trait. (A) shows how the pathogen effect on spVL, $g_v$, evolves according to an OU process along (C) the pathogen phylogeny. The dashed line in (A) shows the expected value of the trait at each point in time under the model. (B) shows that the non-pathogen effect on spVL, $e$, is modeled as a Gaussian random variable. Each tip on the phylogeny corresponds to a sampled individual with an independent non-pathogen-effect drawn from the distribution shown in (B).}
		\label{fig:spVLModel}
	\end{center}
\end{figure}

@@

\begin{figure}[H]
	\begin{center}
		\includegraphics[width=0.75\linewidth]{figures/simulation_results}
		\caption{Results from the simulation study. We simulated pathogen and non-pathogen (host \& environmental) effects on the spVL trait under the POUMM with different trait heritability values (y-axis) and selection strengths (x-axis). For each simulated dataset, we applied our method to estimate the non-pathogen effects and performed GWAS using these estimates. (A) shows how well we can estimate the host effect using our method (left side) compared to just the scaled trait value (right side). (B) shows how GWAS can theoretically be improved by using the true (normally unknown) non-pathogen effects on spVL (left). It can still be improved in certain scenarios using our estimate for this value (middle) compared to the current standard - using just the (scaled) trait value (right). Tile color represents the average value over 20 simulated datasets of 500 samples for each heritability and selection strength value. The black point represents our estimates for the heritability and selection strength of spVL based on Swiss data. RMSE = Root mean square error, TPR = True positive rate.}
		% (RMSE) with respect to simulated values for the host part of spVL ($\bm{{g_h}}$) as a function of spVL trait heritability and selection strength on the trait. The left plot is for the POUMM-based estimate of the non-pathogen part of the trait ($\bm{\hat{e}}$) and the right plot is for the raw trait value scaled by the mean trait value, such that it is centered at zero ($\bm{z} - \bm{\bar{z}}$). (B) True positive rate (TPR) in GWAS performed using true (simulated) values for the non-pathogen part of the trait ($\bm{e}$), the POUMM-based estimate for the non-pathogen part of the trait ($\bm{\hat{e}}$), and the raw trait value scaled by the mean trait value ($\bm{z} - \bm{\bar{z}}$). The color of each plot tile in (A) and (B) represents the average value over 20 simulated datasets of 500 samples. The black point shows our estimates for the heritability of and selection strength on the spVL trait based on Swiss data.}
		\label{fig:simulationResults}
	\end{center}
\end{figure}
